\chapter{Example Applications}
This chapter lists some applications that have been developed using CTL.

\section{Video Doc}
The application can extract video summaries.
\lstinputlisting[caption={Video Doc App}, label=App:VideoDoc, language=ctl]{../experiment/exp1-video/CTL/videoDoc.ctdl}

\section{Journalist}
The application collects news information in the city.
\lstinputlisting[caption={Journalist App}, label=App:Journalist, language=ctl]{../experiment/exp2-Jouranlist/CTDL/Journalist.CTDL}

\section{Map3D}
The application collects 3D coordinate information data in a specified area for building 3D maps.
\lstinputlisting[caption={Map3D App}, label=App:Map3D, language=ctl]{../experiment/exp3-Map3D/CTDL/Map3D.CTDL}

\section{Broken Streetlight}
The app monitors damaged streetlights in the city and assigns crews to fix them, and records the results.
\lstinputlisting[caption={Broken Streetlight App}, label=App:BrokenStreetlight, language=ctl]{../experiment/exp4-SmartCities/CTDL/BrokenStreetlight.CTDL}

\section{Missing Manhole}
The app monitors lost manhole covers in the city and assigns people to fix them, and records the results.
\lstinputlisting[caption={Missing Manhole App}, label=App:MissingManhole, language=ctl]{../experiment/exp4-SmartCities/CTDL/MissingManhole.CTDL}

\section{Rebalance Bicycle}
The app assigns workers to rebalance the distribution of shared bicycles with vehicles.
\lstinputlisting[caption={Rebalance Bicycle App}, label=App:RebalanceBicycle, language=ctl]{../experiment/exp4-SmartCities/CTDL/RebalanceBicycle.CTDL}

\section{Busy Canteen}
The app monitors the flow of customers in the canteen at noon.
\lstinputlisting[caption={Busy Canteen App}, label=App:BusyCanteen, language=ctl]{../experiment/exp5-SmartCampus/CTDL/BusyCanteen.CTDL}

\section{Noise Detection}
The app monitors noise on campus.
\lstinputlisting[caption={Noise Detection App}, label=App:NoiseDetection, language=ctl]{../experiment/exp5-SmartCampus/CTDL/NoiseDetection.CTDL}
